---
title: "Getting Started With qzwsmix99rt"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started With qzwsmix99rt}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(qzwsmix99rt)
devtools::load_all()
```

# Disclaimer
This package contains a collection of R-functions useful to run MiX99 from within R. The pre-requisite for this is that MiX99 is installed in a path that is reachable via the path environment variable.


# First Example
This package contains an example with which the functionality can be tested. The following commands can be used to run the example analysis.

```{r}
test_dir <- prepare_example_p1()
```

This copies the required files for the analysis into a temporary directory. The following shows a list of these files

```{r}
list.files(path = test_dir)
```

The analysis is run with 

```{r}
s_log_file = file.path(test_dir, 
                       paste0(format(Sys.time(), "%Y%m%d%H%M%S"),
                              "_mix99.log", collapse = ""))
run_mix99(ps_work_dir = test_dir, ps_log_file = s_log_file)
```

The results and a logfile are written to the working directory

```{r}
list.files(path = test_dir)
```

The content of the logfile can be inspected by

```{r}
con_log <- file(description = s_log_file)
vec_log <- readLines(con = con_log)
close(con = con_log)
head(vec_log, 10)
```

The end of the logfile

```{r}
tail(vec_log)
```


After the analysis all results can be removed by

```{r}
fs::dir_delete(test_dir)
```


# Conversion of QMSim Output to MiX99 Input
Given an output file from QMSim, the input files for MiX99 are generated. This is done by a conversion function. Currently this works only for simple animal models.

```{r}
s_qm_path <- system.file("extdata", "qmsim", "p1_data_001.txt", package = "qzwsmix99rt")
(s_tmp_dir <- tempdir())
s_out_dir <- file.path(s_tmp_dir, 'mix99_work')
s_file_stem = paste0(format(Sys.time(), "%Y%m%d%H%M%S"),
                                                         "_mix99", collapse = "")
convert_qmsim_to_mix99(ps_qm_path = s_qm_path, 
                       ps_out_dir = s_out_dir, 
                       ps_file_stem = s_file_stem,
                       pl_vcmp      = list(genetic_variance = 0.1, residual_variance = 0.9))
list.files(path = s_out_dir)
```

The content of the generated data file can be inspected by

```{r}
s_data_path <- file.path(s_out_dir, paste(s_file_stem, ".dat", sep = ""))
con_data <- file(s_data_path)
vec_data <- readLines(con = con_data)
close(con = con_data)
head(vec_data)
tail(vec_data)
```


The content of the pedigree can be viewed by

```{r}
s_ped_path <- file.path(s_out_dir, paste(s_file_stem, ".ped", sep = ""))
con_ped <- file(s_ped_path)
vec_ped <- readLines(con = con_ped)
close(con = con_ped)
head(vec_ped)
tail(vec_ped)
```

The content of the parameter file

```{r}
s_var_path <- file.path(s_out_dir, paste(s_file_stem, ".var", sep = ""))
con_var <- file(s_var_path)
vec_var <- readLines(con = con_var)
close(con = con_var)
cat(paste0(vec_var, collapse = "\n"), "\n")
```

The clim file

```{r}
s_clm_path <- file.path(s_out_dir, paste(s_file_stem, ".clm", sep = ""))
con_clm <- file(s_clm_path)
vec_clm <- readLines(con = con_clm)
close(con = con_clm)
cat(paste0(vec_clm, collapse = "\n"), "\n")
```


# Running MiX99 on Prepared Data
The data prepared so far are now used to run mix99

```{r}
s_mix99_log_path <- file.path(s_out_dir,
                              paste0(format(Sys.time(), "%Y%m%d%H%M%S"),
                                     "_mix99.log", collapse = ""))
run_mix99(ps_work_dir = s_out_dir, ps_log_file = s_mix99_log_path)
```

The content of the working directory

```{r}
list.files(path = s_out_dir)
```

The log file is

```{r}
con_mix99_log <- file(s_mix99_log_path)
vec_mix99_log <- readLines(con = con_mix99_log)
close(con = con_mix99_log)
cat(paste0(vec_mix99_log, collapse = "\n"), "\n")
```


# Problems with Recoding NA Values

```{r}
tbl_qm <- read_qmsim(ps_path = s_qm_path)
```

```{r}
(tbl_data <- dplyr::select(tbl_qm, Progeny, Sex, Phen))
```

```{r}
(tbl_data <- dplyr::mutate(tbl_data, Sex = dplyr::recode(Sex, `M` = 1, `F` = 2)))
```

```{r}
pn_mvc       = -99999.0
(tbl_data <- dplyr::mutate(tbl_data, Phen = tidyr::replace_na(Phen, pn_mvc)))

```

# Clean Up
The generated files are removed by 

```{r}
fs::dir_delete(s_tmp_dir)
```

